---
title: 'Part 1: binary logistic regression'
author: "Grace Brofman"
date: "1/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(palmerpenguins)
library(GGally) # for pairs plots
library(broom) # for tidy model outputs in dataframes
library(jtools) # to make quick graphs of model outputs
```

```{r}

penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color = species)) # to make a pairs plot

# we're looking specifically at adelie and chinstrap bc they tend to overlap in features
# make subset only containing adelie and chinstrap (see levels(penguins$species))
```

create subset with only adelie and chinstrap
```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c("Adelie", "Chinstrap")) %>% # agian, check (see levels(penguins$species)) and see that "Gentoo" still shows up as a factor level
#factor_drop() will help us get rid of that
  mutate(species = fct_drop(species)) #drops any factor levels for which there are no acutal observationa

```

Let's just do a dataviz of those 2 species:
```{r}
ggplot(data = adelie_chinstrap, 
       aes(x = body_mass_g,
           y = flipper_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)

# see sexual dimorphism, looks like chinstraps tend to be a bit larger in flipper length
# should look at more visualizations to understand differences between the species
# can binomial logistic regression help us understand the relationship between several predicotr variables and the probablillty based on those variables that a penguin is adelie or chinstrap?
```

Binary logistic regression
- body mass
- flipper length
- penguin sex

```{r}
adelie_chinstrap_blr <- glm(species ~ body_mass_g + flipper_length_mm + sex,#generalized linear model, species is the dependent variable
                            data = adelie_chinstrap,
                            family = "binomial") # what is that link on the left hand side? (log odds in this case)

```

Take a look at the outcomes:
```{r}
adelie_chinstrap_blr #tells intercept value and coefficients (here we see that female is reference level for sex bc it didnt show up)

summary(adelie_chinstrap_blr) # gives more info (see that p-values show us that of the 3 variables, only flipper length is significantly non-zero- although doesnt necessarily mean we should drop the other variables)

#take body mass coefficient( -8.6 x 10^-4)
# type levels(adelie_chinstrap$species) into console to see reference levels, we see that Adelie is listed first- this means that Adelie is coded as 0 and chinstrap is coded as 1. so in our log linear expression, the left-hand side will be the probablity of chinstrap (the log odds of the non-zero level)
# we expect on average that as body mass increases, the odds of it being a chinstrap are going down (the sign indicates this), although it's a very small coefficient compared to others (p value is nonsignificant)
# flipper length coefficient - positive value indicates that increase in length means odds of chinstrap increases
# sexmale: technically interpreted as compared to female level, if a penguin is male its less likely to be achinstrap (in this case its a bit nonsensical). also the p value is nonsignificant
# does it make sense that as flipper lenght increase, the odds of being a chinstrap increase too? 
# do some data viz!
```

```{r}
ggplot(data = adelie_chinstrap,
       aes(x = species,
           y = flipper_length_mm)) +
  geom_jitter(aes(color = sex))
# yes, maybe chinstrap flipper length gets larger, it looks a little more likely that you're a chinstrap 
```
look at actual probabioilities using broom::augment()
What are the actual probabilities of being Chinstrap for each of the existing observations in our adelie_chinstrap data frame?
takes original data, adds a column with the probabilites that it is a chinstrap penguin
```{r}
blr_fitted <- adelie_chinstrap_blr %>% 
  broom::augment(type.predict = "response") 

# now added on to our original df, we can see a bunch of info added for our model outcomes
# .fitted column is based on our blr, the prob based on 3 predictor variables, that our penguin is a chinstrap. we see that for most adelies, there's a less than 50% chance that it's a chinstrap penguin. probabilities tend to be higher that its a chinstrap when we're looking at chinstraps, but it still incorrectly characterizes them sometimes

```

Plotting fitted probabilities:
```{r}
ggplot(data = blr_fitted,
       aes(x = flipper_length_mm,
           y = .fitted)) +
  geom_point(aes(color = sex, pch = species)) +
  geom_smooth(aes(color = sex), se = FALSE) + # se refers to error ribbon
  labs(x = "Flipper length (mm)",
       y = "Probability of outcome Chinstrap")

# we see an increasing probablility of chinstrap with increasing flipper length. we also see the diff based on sex
# then we add point classifier for species. we thensee that there are plenty of adelies that are given higher than 50% chance of being a chinstrap. our model is not perfect

```

Plot model based on different variables:
```{r}
effect_plot(adelie_chinstrap_blr,
            pred = flipper_length_mm, # pred = predictor variable
            y.label = "Probability of Chistrap",
            interval = TRUE)

effect_plot(adelie_chinstrap_blr,
            pred = body_mass_g, # pred = predictor variable
            y.label = "Probability of Chistrap",
            interval = TRUE)

# effect plots are not as customizable, but give quick overviews of how your probability changes with respect to a model variable

```

What's the probability that a female penguin weighing 3410 g with a flipper length of 192 mm is a Chinstrap?

```{r}
ex_1 <- predict(adelie_chinstrap_blr,
                data.frame(sex = "female",
                           body_mass_g = 3410,
                           flipper_length_mm = 192),#give it a df containing the values that you want to calculate this for (ours has 1 row but you can feed it multiple rows). we also have to make the variables match those for which the model was built e.g. "female", the reference level
                type = "response") 

ex_1 # this returns 0.402 which means that using the model we've created based on these characteristics of this hypothetical new penguin, there's a 40% chance that its a chinstrap penguin

```





